---
title: "Decision and Boosted Tree"
author: "Zhengkan Yang"
date: "Tuesday, August 12, 2014"
output: html_document
---

# Decision and Boosted tree modeling for predicting repairs

## Outline
* Read the data and load the library
* Decision Tree
* Boosted Tree


## Install the necessary package dependencies
```{r include = FALSE}
# install.packages("gbm")
# install.packages("caret")
# install.packages("rpart")
# install.packages("corrgram")
# install.packages("party")
# install.packages("tree")
# install.packages("MESS")
# install.packages("rpart.plot")
```


## Step 1: Read the data, load the library and split the data set
**Note**: Please change the data_name here to your own version of data 
```{r include=FALSE}
library(gbm)
library(caret)
library(rpart)
library(corrgram)
library(party)
library(tree)
library(MESS)
library(rpart.plot)
```

```{r}
data_name = "C:/Users/zhengkya/Desktop/Crash&Repair Data/CSV/Mapping_recent_crash_repair.csv"
data <- read.csv(data_name,header = TRUE)
```

### Split the data set into training and text
```{r}
row_train <- sample(seq(1,nrow(data),by=1),0.6*nrow(data),replace=FALSE)
data_train <- data[row_train,]
data_test <- data[-row_train,]
```



## Step 2: Classification Tree Model
Package used: `rpart`

Evaluation Metric: Use test data set to get the **Receiver Operator Curve** and compute the Area Under Curve (**AUC**)

###Fit a tree on training set
```{r}
data_train$repair_id[which(data_train$repair_id == 1)] <- "repaired"
data_train$repair_id[which(data_train$repair_id == 0)] <- "Norepair"
control = tree.control(nobs=nrow(data_train), mincut = 20, minsize = 40, mindev = 0.0005)
tree.fit <- rpart(repair_id~.,data_train,control=rpart.control(minsplit = 20,minbucket = 40,cp=0.01),method="class")
```


###Visualized the best tree
```{r}
plot(tree.fit)
text(tree.fit, use.n=TRUE)
prp(tree.fit, extra=6, uniform=T, branch=0.5, yesno=T, border.col=1, type=4,xsep="/",box.col="green",branch.col="red")
```

###Score the test set use the fitted model and make ROC plots 
```{r}

#Predict true values/false values
pred <- predict(tree.fit, newdata=data_test,type="prob")[,2]
obs = data_test$repair_id

#Calculate ROC
threshold = 0.5
roc.curve= function(threshold,print = FALSE){
  Ps = (pred>threshold)*1
  FP = sum((Ps==1)*(obs==0))/sum(obs==0)
  TP = sum((Ps==1)*(obs == 1))/sum(obs==1)
  if(print==TRUE){
    print(table(Observed = obs,Predicted=Ps))
  }
  vect = c(FP,TP)
  names(vect) = c("FPR","TPR")
  return(vect)
}
roc.curve(threshold,print=TRUE)

ROC.curve <- Vectorize(roc.curve)
Grid.ROC <- ROC.curve(seq(0,1,by=0.01))

plot(Grid.ROC[1,],Grid.ROC[2,], lwd = 3, col = "red", xlab = "False Positive Rate", ylab = "True Positive Rate",lty=2)

seg_mat <- Grid.ROC[,c(1,16,47,98)]

for(i in c(1:4)){
  if(i!=4){
    segments(seg_mat[1,i],seg_mat[2,i],seg_mat[1,i+1],seg_mat[2,i+1],col="red",lwd = 3,lty=1)
  }
}
lines(seq(0,1,by=0.01),seq(0,1,by=0.01),col= "blue",lwd=3)

print(paste("The area under curve is: ", toString(auc(Grid.ROC[1,],Grid.ROC[2,],type = 'spline')),sep=""))

```


## Step 3: Boosted Tree Model
* Package Used: `caret` and `gbm`
* Evaluation metric: ROC and AUC
### Fit the model use caret
```{r eval=FALSE}
fitControl <- trainControl(method="repeatedcv",number=10, repeats=1,classProbs = TRUE,
                            summaryFunction = twoClassSummary)
tuneGrid = expand.grid(n.trees = c(100,200,300), interaction.depth=c(3,4,5),shrinkage=c(0.05,0.1,0.15))
boostedTr <- train(as.factor(repair_id)~.,data=data_train,tuneGrid = tuneGrid, trControl=fitControl,method="gbm")

```


###The best boosted tree model
* Number of trees: 300
* Interaction.depth: 5
* Shrinkage: 8
```{r}
data_train$repair_id[which(data_train$repair_id == "repaired")] <- 1
data_train$repair_id[which(data_train$repair_id =="Norepair")] <- 0
boostedTr <- gbm(repair_id~.,data=data_train,distribution="bernoulli", n.trees = 300, interaction.depth= 5, shrinkage=0.1)

pred <- predict(boostedTr, newdata=data_test,n.trees=300,type="response")
obs <- data_test$repair_id
```


###Plot relative importance for each predictor
```{r}
x <- data.frame(summary(boostedTr),row.names=NULL)
barplot(x$rel.inf, beside=T, xlab = "Bucket Area", ylab = "Relative Importance", main="Boosted Tree Variable Importance",horiz=FALSE, names.arg = row.names(x),las=1, cex.names=0.7,col="blue")
```



###Plot ROC and get AUC
```{r}
threshold = 0.5
roc.curve= function(threshold,print = FALSE){
  Ps = (pred>threshold)*1
  FP = sum((Ps==1)*(obs==0))/sum(obs==0)
  TP = sum((Ps==1)*(obs == 1))/sum(obs==1)
  if(print==TRUE){
    print(table(Observed = obs,Predicted=Ps))
  }
  vect = c(FP,TP)
  names(vect) = c("FPR","TPR")
  return(vect)
}
roc.curve(threshold,print=TRUE)

ROC.curve = Vectorize(roc.curve)

M.ROC=ROC.curve(seq(0,1,by=.01))
plot(M.ROC[1,],M.ROC[2,],lwd=3,col="red",type="l",xlab = "False Positive Rate", ylab="True Positive Rate")

lines(seq(0,1,by=0.01),seq(0,1,by=0.01),col="blue",lwd=3)


print(paste("The area under curve is: ", toString(auc(Grid.ROC[1,],Grid.ROC[2,],type = 'spline')),sep=""))

```



